--- 
title: "The Association Between Travel Behavior and Urban Form"
author: 
- Shen Qu^["Portland State University, Toulan School of Urban Studies and Planning"]
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: 
  bookdown::gitbook: 
    lib_dir: "book_assets"
    includes:
      after_body: disqus.html
  bookdown::pdf_book:
    toc: true
    toc_depth: 4
  citation_package: biblatex
documentclass: book
bibliography: [fieldpaper0822.bib, unsorted.bib, book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: qushen26/field_paper
description: "This is a field paper using the bookdown package. The output format for this example is bookdown::gitbook."
# abstract: |

  
keywords: travel behavior, urban form, regression analysis
header-includes:
   - \usepackage{amssymb}
   - \usepackage{amsmath}
   - \usepackage{fancyhdr}
   - \pagestyle{fancy}
   # - \fancyhead[CO,CE]{\thesection}
   # - \fancyfoot[CO,CE]{\thesection}
   # - \fancyfoot[LE,RO]{\thepage}
   # - \fancyhf{}
   # - \rhead{\thechapter}
   # - \chead{Field Paper}
   # - \lhead{\thesection}
   # - \rfoot{Page \thepage}
   - \usepackage{setspace}\doublespacing
   - \usepackage{subfig}
   - \usepackage{multicol}
   - \usepackage{multirow}


---

# Preface {-}

This field paper is for discussing the relationship between travel behavior and urban form. ^[This is a field paper using the **bookdown** package [@R-bookdown], **knitr** [@xie2015] and TinyTeX: <https://yihui.name/tinytex/>.]
Part I reviews the related literature and tries to cover the main theories and research in this field.
Daily Driving Distance as the variable of interest, is associated with many factors including urban form variables.
The discussion will gradually narrow down to the association between driving distance and urban density. 
Instead of using very high dimension model or synthesized index, 
Population Density itself may have more potential to explain the driving distance than people ever thought about.
  
Part II introduces the linear, generalized linear, and non-linear models which is the fundamental methods for the question raised in Part I.
Meta-analysis is also a useful method to get more general results based on many studies with the same topic.
It can be applied on investigating self-selection and publication bias.



```{r setup, include=FALSE}
# knitr::opts_chunk$set(
#   tidy=FALSE,     # display code as typed
#   size="small")   # slightly smaller font for code
# devtools::install_github('rich-iannone/DiagrammeR')
# remotes::update_packages(c("DiagrammeR", "tinytex", "rmarkdown", "knitr"))
# remotes::install_github('rstudio/rmarkdown')
library(pacman)
p_load(tidyverse,kableExtra ,mosaic,ggpubr,factoextra, DiagrammeR,magrittr) #
# if (!require("webshot")) {install.packages("webshot")}
# webshot::install_phantomjs()

knitr::opts_chunk$set(eval=F,echo=F, message=F, warning=F,cache=T,fig.dim = c(8, 6),fig.align = 'center') #,fig.pos="!h"
```


```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
bookdown::publish_book(account="")
```


```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```


