---
title: "The Association Between Travel and Urban Form"
subtitle: "Part I First field: Theories and Framework"
author: "Shen Qu"
date: "`r Sys.Date()`"
# site: bookdown::bookdown_site
output: 
  # bookdown::gitbook: 
  #   lib_dir: "book_assets"
  #   includes:
  #     after_body: disqus.html
  bookdown::pdf_book:
    toc: true
    toc_depth: 4
  citation_package: biblatex

documentclass: article #ElegantPaper # memoir # custom # amsbook  # scrbook # book
bibliography: [fieldpaper1114.bib, unsorted.bib] #, book.bib, packages.bib
biblio-style: apalike
link-citations: yes
# github-repo: qushen26/field_paper
description: "This is a field paper using the bookdown package."
# abstract: |
always_allow_html: true
fontsize: 12pt
geometry: margin=1.25in
# classoption: openany # remove blank pages
keywords: travel behavior, urban form, regression analysis
header-includes:
   - \usepackage{amssymb}
   - \usepackage{amsmath}
   - \usepackage{fancyhdr}
   - \pagestyle{fancy}
   - \setlength{\headheight}{15pt}
   # - \fancyhead[CO]{\thesection}
   # - \fancyhead[CE]{\thesection}
   # - \fancyfoot[CO,CE]{\thesection}
   # - \fancyfoot[LE,RO]{\thepage}
   # - \fancyhf{}
   # - \rhead{\thechapter}
   # - \chead{Field Paper}
   # - \lhead{\thesection}
   # - \rfoot{Page \thepage}
   # - \usepackage{setspace}\doublespacing
   - \usepackage{subfig}
   - \usepackage{multicol}
   - \usepackage{multirow}
---

# Preface {.unnumbered}

This field paper is for discussing the relationship between travel and urban form. The initial motivation is curiosity about how the travel distance is affected by urban densities. Is this relationship existing universally or just context-dependent? Are these models in literature replicable or reproducible? Part I reviews the related literature and tries to cover the main theories and research in this field. Travel patterns or behaviors as the variable of interest are associated with many factors, including urban form variables. To answer the above questions, this field paper extends to different types of travel variables and more influencing factors.

The travel-urban form studies are many and perplexing. Part II introduces the common statistical methods for the question raised in Part I. Although emerging many new techniques in recent years, regression analysis is still the fundamental and explainable methodology. The first stage is to understand the different methods adopted in previous studies. After that, it could be possible to give fair criticism or replicate the previous models.

Under the travel-urban form topic, one goal is to examine the validation of different models. A parallel study on known data may find the issues of mismatch or misreading and improve the relevant studies in the future. Another potential goal is to get some generalized association between travel and urban form. Based on many studies on the same topic, a meta-analysis is a useful tool that can help to disclose publication bias in this field.

<!-- - Several Ideas -->

<!-- The discussion will gradually narrow down to the association between driving distance and urban density.  -->

<!-- Instead of using a very high dimension model or synthesized index,  -->

<!-- Population Density itself may have more potential to explain the driving distance than people ever thought about. -->

<!--  Many disaggregated travel-urban form models without psychological factors are underfitting, the corresponding estimates are biased. -->

<!-- When the goal of related studies is to provide evidence for public policy, the aggregated analysis may be good enough. -->

<!--  The estimated coefficients are incomparable unless the regression models have the common key specification. Because $\boldsymbol{\hat\beta}=(\mathbf{X'X})^{-1}\mathbf{X'}\mathbf{y}$ is a function depending on $\mathbf{X}$. -->

<!--  Standardized coefficients or elasticities cannot fix these issues. -->

<!--  Due to the non-linear characteristics of many individual and urban-form factors, identifying the effective ranges of variables is useful for policymaking. -->

```{r setup, include=FALSE}
# knitr::opts_chunk$set(
#   tidy=FALSE,     # display code as typed
#   size="small")   # slightly smaller font for code
# devtools::install_github('rich-iannone/DiagrammeR')
# remotes::update_packages(c("DiagrammeR", "tinytex", "rmarkdown", "knitr"))
# remotes::install_github('rstudio/rmarkdown')
library(pacman)
p_load(tidyverse,kableExtra ,mosaic,ggpubr,factoextra, DiagrammeR,magrittr) #
# if (!require("webshot")) {install.packages("webshot")}
# webshot::install_phantomjs()
knitr::opts_knit$set(root.dir = "C:/E/PhD/courses/Field paper")
knitr::opts_chunk$set(eval=T,echo=F, message=F, warning=F,cache=T,fig.align = 'center') #,fig.pos="!h",fig.dim = c(8, 6)

```

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
bookdown::render_book("index1.Rmd", bookdown::pdf_book(keep_tex = TRUE))
# bookdown::render_book("index.Rmd", bookdown::word_document2())
# bookdown::render_book()
# bookdown::publish_book(account="qushen")
```

```{r include=FALSE,eval=F}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
