<!-- # (PART\*) Part II Statistical Methods {-}  -->

<!-- {.unnumbered}  -->

<!-- # Model Specification -->

# Modeling Steps

Regression analysis is still the dominant method for explaining the relationship between urban form and travel. When we fit a regression model to connect both sides - predictors, and response - the types of travel variables determine the model's forms. Continuous, count and categorical variables have distinctive characteristics. Correctly choosing a genre of the model is the first step of modeling.

<!-- as below: -->

<!-- \begin{equation} -->

<!-- \mathbf{Y}=\mathbf{X}\boldsymbol{\beta}+\boldsymbol{\varepsilon} -->

<!-- (\#eq:lm) -->

<!-- \end{equation} -->

<!-- where $\mathbf{Y}$ are the variable of VMT. $\mathbf{X}$ are all relevant covariates with corresponding coefficients $\boldsymbol{\beta}$. $\boldsymbol{\varepsilon}$ are a random error term with expected value of $0$ and variance $\sigma^2$. -->

## Model forms

### Logarithm Transformation

<!-- ## For Trip Distance -->

In travel-urban form studies, travel distance belongs to this category. Other continuous variables, such as travel time, CO$_2$ emission, are also usual research interests in transportation. When choosing a continuous response, like Vehicle Miles Traveled (VMT) [^methods-1], the basic structure of the regression model is the response variable equals the linear combination of coefficients and covariates plus a random error term with zero mean and variance. Most VMT-urban form models use this framework [@stevensDoesCompactDevelopment2017; @gimRelationshipsLandUse2013].

[^methods-1]: or Vehicle Kilometers Traveled (VKT)

But in literature, many regression models take logarithm transforms on all variables, called **log-log** models. A recent example is that @alamFactorsAffectingTravel2018 use a log-log model for travel demand by transit at the MSA level in the U.S. Some only transform one or a part of variables [@bentoImpactUrbanSpatial2003; @perumalContextualDensityUS2017], called **log-linear** models. Meanwhile, some models keep the original metrics of data [@cerveroTravelDemand3Ds1997; @zhouSelfSelectionHomeChoice2008; @zhangWhenContextMeets2020]. Which one is the correct method?

Many studies don't explain their choice and treat log transform as a 'tradition.' Some of them clarify that travel distance is highly right-skewed distributed, log transform can address the issues of normality [@ewingVaryingInfluencesBuilt2015]. Hence, Log transformation on travel distance is widely used in transportation-related research and practice.

Another 'tradition' addresses the zero issue by adding one on travel variable [@ewingVaryingInfluencesBuilt2015]. A few studies choose the **Tobit model** to deal with the domain and normality issues [@chatmanHowDensityMixed2003; @chatmanDeconstructingDevelopmentDensity2008; @boarnetWalkingUrbanDesign2008]. Tobit model assumes the travel distance follows a left-censored normal distribution. That means no log transform is needed, but a specific distribution must be accounted for. Does unobserved negative distance or utility exist? The Tobit model claims a firm assumption and requires theoretical and empirical evidence. Normality is not the requirement for the response variable. The fundamental critical condition is linearity and additivity [@montgomeryIntroductionLinearRegression2021].

<!-- For continuous response variables, Multiple Linear Regression (MLR) is the proper type of model, and Ordinary least squares (OLS) is the corresponding algorithm.  -->

<!-- However, These variables have a common feature: their domain is a positive number, not the whole real number field. Here may raise a debate of whether zero values are a part of these variables or not. -->

<!-- As mentioned in the previous chapter, the logarithm transformation on these variables can convert the positive values to real numbers, and the zero values are excluded automatically.  -->

<!-- [@alamFactorsAffectingTravel2018;@leeComparingImpactsLocal2020] -->

<!-- This question should be answered by checking model adequacy, presented in the next section.  -->

<!-- A correct model type choice may depend on the data and research design. -->

<!-- Obviously, the models with and without log transform have different structures and are not equivalent. -->

<!-- The following question is, do their outcomes are comparable? -->

<!-- This question needs further investigation. -->

<!-- If the answer is no, the relevant meta-analysis or summaries should tread them separately. -->

<!-- - Which approach is the proper one?  -->

<!-- Here gives two functions of log transform from a mathematical perspective: variance stabilizing and linearizing. -->

<!-- - Variance Stabilizing -->

<!-- Equality of variance is a primary assumption of the regression model.  -->

<!-- When variance is not constant, the least-squares estimators will not give the minimized variance. -->

<!-- Though the estimation is still unbiased, the standard errors of regression coefficients will be larger, and the model becomes insensitive. -->

<!-- @montgomeryIntroductionLinearRegression2021 give several useful variance stabilizing transformations  -->

<!-- in Table @ref(tab:transformation). -->

```{r,eval=F}
library(kableExtra) 
kbl(data.frame(
  `Relationship`=c('$\\sigma^2\\propto E[y]$','$\\sigma^2\\propto (E[y])^2$','$\\sigma^2\\propto (E[y])^3$','$\\sigma^2\\propto (E[y])^4$'),
  `Transformation` = c('$y^{1/2}$','$\\ln(y)$','$y^{-1/2}$', '$y^{-1}$')
)  , booktabs = TRUE, label = 'transformation', digit=2, escape = F, #, align = "llr"
  caption = 'variance stabilizing transformations'
) %>% kable_classic(full_width = F,font_size = 7) 
```

<!-- A preliminary study using National Household Travel Survey (NHTS) [@nhts_2009] data finds both the mean and standard deviation of household daily VMT are close to 40. -->

<!-- This relationship supports that the logarithm of $\mathbf{y}$ is a proper choice for variance stabilizing. -->

<!-- - Linearizing -->

<!-- Another fundamental assumption, linearity is also can be addressed by transformation. -->

<!-- If the relationship between response and predictors is linearizable, a suitable transformation can construct a intrinsically linear model. -->

<!-- Several common forms from [@montgomeryIntroductionLinearRegression2021] are  -->

<!-- shown in Table @ref(tab:linearlizing). -->

```{r,eval=F}
library(kableExtra) 
kbl(data.frame(
  `Relationship`=c('$y=\\beta_0\\exp[\\beta_1x]\\varepsilon$',
                   '$y=\\beta_0+\\beta_1\\ln(x)+\\varepsilon$',
                   '$y=\\beta_0x^{\\beta_1}\\varepsilon$',
                   '$y=x/((\\beta_0+\\varepsilon)x+\\beta_1)$'),
  `Transformation` = c('$y^*=\\ln(y),\\varepsilon^*=\\ln\\varepsilon$','$x^*=\\ln(x)$','$y^*=\\ln(y),x^*=\\ln(x),\\varepsilon^*=\\ln\\varepsilon$', '$y^*=1/y,x^*=1/x$'),
  `Linear Form` = c('$y^*=\\ln \\beta_0 +\\beta_1x +\\varepsilon^*$',
                    '$y=\\beta_0 +\\beta_1x^*+\\varepsilon$',
                    '$y^*=\\ln\\beta_0 +\\beta_1x^* +\\varepsilon^*$',
                    '$y^*=\\beta_0 +\\beta_1x^* +\\varepsilon$')
)  , booktabs = TRUE, label = 'linearlizing', digit=2, escape = F, #, align = "llr"
  caption = 'Linearizing transformations'
) %>% kable_classic(full_width = F,font_size = 7) 
```

<!-- Comparing these forms, the $log(y)$ transformation also called log-linear model gives a finite value of response $y$ when predictor $x\to  0$.  -->

<!-- While the log-log model ($y'=\ln(y),x'=\ln(x)$) will give an infinite value of $y$ when $x\to  0$.  -->

<!-- This gives a useful hint when one chooses from log-linear and log-log models. -->

<!-- Moreover, the $log(y)$ transformation changes the scale of error term. -->

<!-- Only one term in $\varepsilon$ and $\ln\varepsilon$ can be close to constant mean and normal distributed. -->

<!-- Therefore, residual diagnosis is still a effective way for choosing the proper form of transformation. -->

<!-- Prior theories and experience can also help to make a proper choice. -->

In Part I, we discuss the mechanism affecting travel response and the properties of travel variables. Recalling the equation of gravity model, both Gravity law and Zipf's law imply that a logarithm transformation on VMT is suitable.

### Count Data Models

The frequency of trips or ridership is a count variable. <!--  The observed counts of trips $Y_i,...,Y_n$ are random variables aggregated over differing numbers of individual or household with support $Y=\in\{0,1,2,...\}$. --> The trips as events occur randomly in a day or other time. A usual assumption is that count data follow a Poisson or negative binomial distribution. That means their link function is the logarithm. It looks like the logarithm on response is the same as the log-linear model. Which one is correct?

@choiAnalysisMetroRidership2012 take log transform on both sides of the equation and compare the performance between Poisson regression and log-log models (called multiplicative models). They think "the Poisson model ... reflects the varying elasticity of the dependent variable according to the level of independent variables." @choiAnalysisMetroRidership2012 purpose log-log model is better for greater F-statistic and adjusted $R^2$ than the Poisson model. But the test statistic is not a good measurement for comparing models with different structures. $R^2$ is only a piece of evidence for the goodness of fit.

Using Monte Carlo simulation, @wangDetransformationBiasNonlinear2018 's study shows that the predictions of the log-transform model may have a significant bias when conducting detransformation. The comparison between log-transform and Poisson models needs to go back to the data properties and underlying mechanism. The detransformation bias may be due to the inappropriate model structure. Although the two types of models have similar forms of the equation, they perform two distinct types of randomization. A convincing comparison still calls for adequacy checking.

<!-- There are three conditions for the Poisson process: -->

<!-- First, as a stochastic process, the probability of at least one event happening in a time interval is proportional to the length of the interval. -->

<!-- Second, the probability of two or more events occurring in a small time interval is close to zero. -->

<!-- Third, in disjoint time intervals, the count numbers of the trip should be independent. -->

<!-- In real life, a traveler can not make two trips simultaneously, so the second condition holds. -->

<!-- But a household with two workers and two students might have four trips at the same time every morning. Hence, individual count data is more valid than households' when using Poisson distribution.  -->

<!-- The independence of count number among different time intervals may not be valid too. The daily trips often belong to a trip chain and require more information at a micro-level. -->

Negative binomial regression has the same link function with Poisson models. For the advantage of addressing overdispersion, more travel-urban form studies are choosing negative binomial regression than Poisson models [@ewingVaryingInfluencesBuilt2015]. Some studies found the estimated coefficients are similar in two types of models [@chatmanDeconstructingDevelopmentDensity2008]. @dillPredictingTransitRidership2013 also report that count-data models have no apparent advantages in prediction. Research about interval estimation may disclose their difference. Quasi-Poisson Model is another simple way for overdispersion.

<!-- (Equation @ref(eq:pois-link)) -->

There are often more no-trip observations in trip frequency data than Poisson, and negative binomial distribution is expected. One problem in previous planning studies is manipulating data by replacing zero values with one [@ewingVaryingInfluencesBuilt2015]. But the meaning and mechanism of 'No participate' are essentially different from that of intensive of participating [@greeneEconometricAnalysis2018, 18.4.8]. There are two methods, the Zero-inflated model and hurdle model can address this issue. Both of them assume the data arise from two mechanisms.

@ewingVaryingInfluencesBuilt2015 firstly apply the hurdle models on travel-urban form study. @zhangHouseholdTripGeneration2019 use the Zero-inflated Negative Binomial model to examine the built environment's influences on trip generation. But their articles are not for testing the advantage of the two-step model. Conceptually, the two-step models can better express the decision process discussed in Part I. It is worth comparing the performance between hurdle, Tobit, and replacing-0-with-1 models in the future.

<!-- - Poisson Regression -->

<!-- The probability mass function (pmf) of Poisson distribution and its canonical form can be found in @agresti2015foundations and @rodriguez2007generalized. -->

<!-- So Poisson distribution has a simple link function as -->

<!-- And Poisson distribution has the property of $E[y_i]=Var[y_i]=\mu_i$ as the systematic component. -->

<!-- By taking log transform, the non-negative parameter space mapping to real number. -->

<!-- It also convert the multiplicative relationship among predictors to additive. -->

<!-- The value of coefficient $\beta_j$ means that per unit change in predictor $x_j$ leads to the expected change in the log of the mean of response. -->

<!-- Another interpretation is that the mean of response would multiple $\exp[\beta_j]$ by per unit change in $x_j$. -->

<!-- Iteratively reweighed least squares method (IRLS) can solve the log-linear Poisson model. -->

<!-- The key correction step is $\hat{\eta_i}^{(1)}=\hat{\eta_i}^{(0)} + \frac{y_i-\hat\mu_i^{(0)}}{\hat\mu_i^{(0)}}$. -->

<!-- The diagonal weight matrix is $w_{ii}=\hat\mu_i^{(0)}$ -->

<!-- \begin{equation} -->

<!-- Pr(Y=y) = \frac{e^{-\mu}\mu^y}{y!}=\exp[\log(\mu) y-\mu](y!)^{-1} -->

<!-- \end{equation} -->

<!-- \begin{equation} -->

<!-- \begin{split} -->

<!-- g(\mu_i)&=\log\mu_i=\eta_i=\mathbf{x}'\boldsymbol{\beta}\\ -->

<!-- g^{-1}(\eta_i)&=\exp[\eta_i]=\mu_i=\exp[\mathbf{x}'\boldsymbol{\beta}]\\ -->

<!-- \end{split} -->

<!-- (\#eq:pois-link) -->

<!-- \end{equation} -->

<!-- - Negative Binomial Regression -->

<!-- A restriction of Poisson Distribution is that the mean and variance should be equal or proportional. -->

<!-- In many count data, the inequality of them is called overdispersion. -->

<!-- Overdispersion is very common in trip frequency data. It could be because the daily trips are not independent and homogeneous. They are a mixture of several different purposes or several persons in a household. -->

<!-- By adding a new parameter, mixture rate can construct a Poisson mixture model and address the overdispersion. -->

<!-- Suppose an unobserved random variable follow a gamma distribution $Z\sim Gamma(r,1/r)$ where $r$ is the shape parameter. The pdf can be found in @agresti2015foundations and @rodriguez2007generalized. -->

<!-- \begin{equation} -->

<!-- f(z)=\frac{r^r}{\Gamma(r)}z^{r-1}\exp[-rz],\quad z>0 -->

<!-- \end{equation} -->

<!-- It has $E[Z]=1$ and $Var[Z]=1/r$.  -->

<!-- Then a mixture model can be denote as a conditional distribution $Y|Z\sim Pois(\mu Z)$ for some $\mu>0$ and -->

<!-- $E[Y]=\mu$ and $Var[Y]=\mu+\frac{\mu^2}{r}$. -->

<!-- It is called Poisson-Gamma distribution who can represents the inequality of mean and variance. -->

<!-- If $r$ represent the given number of success and $y$ represent the observed number of failure in a sequence of independent Bernoulli trails. Then the success probability  is $p=r/(r+\mu)$ -->

<!-- Recall that $\Gamma(r+y)=\int_0^\infty z^{r+y-1}\exp[-z]dz$, it can be proved that $Y$ follow a negative binomial distribution. -->

<!-- \begin{equation} -->

<!-- E[Y]=E[E[Y|Z]]=E[\mu Z]=\mu E[Z]=\mu -->

<!-- \end{equation} -->

<!-- and -->

<!-- \begin{equation} -->

<!-- \begin{split} -->

<!-- Var[Y]&=E[Var[Y|Z]] + Var[E[Y|Z]]\\ -->

<!-- &=E[\mu Z]+Var[\mu Z]\\ -->

<!-- &=\mu E[Z] + \mu^2Var[Z]\\ -->

<!-- &=\mu+\frac{\mu^2}{r} -->

<!-- \end{split} -->

<!-- \end{equation} -->

<!-- \begin{equation} -->

<!-- \begin{split} -->

<!-- p(y)&=\int_0^\infty p(y|z)\cdot f(z)dz\\ -->

<!-- &=\int_0^\infty \frac{(\mu z)^y\exp[-\mu z]}{y!}\cdot\frac{r^r}{\Gamma(r)}z^{r-1}\exp[-rz]dz\\ -->

<!-- &=\frac{r^r\mu^y}{y!\Gamma(r)}(\mu+r)^{-y-r}\int_0^\infty [(\mu+r)z]^{y+r-1}\exp[-(\mu+r)z]d(\mu+r)z\\ -->

<!-- &=\frac{\Gamma(r+y-1)}{\Gamma(y+1)\Gamma(r)}(\frac{r}{\mu+r})^r(\frac{\mu}{\mu+r})^y\\ -->

<!-- &={{r+y-1}\choose{r-1}}p^r(1-p)^y, \quad y=0,1,2,... -->

<!-- \end{split} -->

<!-- \end{equation} -->

<!-- However, negative binomial distribution is non-exponential families. Using maximum likelihood method and log link, the coefficients can be estimated. -->

<!-- Quasi-Poisson Model is another simple way for overdispersion.  -->

<!-- It introduces a dispersion parameter $\phi$. The Poisson model has $Var[Y|\eta]=\phi\mu$ where $\phi>1$.  -->

<!-- The estimated $\phi$ is $\hat\phi=\frac{1}{n-p}\sum\frac{(Y_i-\hat\mu_i)^2}{\hat\mu_i}$ -->

<!-- The extra parameter can  be estimated by maximum likelihood. -->

<!-- - Zero-inflated and Hurdle Models -->

<!-- In trip frequency data, there are often more no-trip observations than Poisson and negative binomial distribution expected. One problem in previous planning studies is to manipulate data by replacing zero values with one [@ewingVaryingInfluencesBuilt2015]. But the meaning and mechanism of 'No participate' are essentially different with that of intensive of participate [@greeneEconometricAnalysis2018, 18.4.8]. -->

<!-- Zero-inflated model and hurdle model can address this issue. -->

<!-- Both of them assume the data arise from two mechanisms. -->

<!-- For example, @zhangHouseholdTripGeneration2019 use the Zero-inflated Negative Binomial model to examine the influences of built environment on trip generation. -->

<!-- In Zero-inflated Poisson/negative binomial model, both of two mechanisms generate zero observations.   -->

<!-- The first mechanism produce the excess zeros with $\pi_i=Pr(Y_i=0)$. The rest zero and positive values are generated by the second mechanism $f(y;\mathbf{x}_i,\boldsymbol\beta)$ of Poisson or negative binomial pmfs. -->

<!-- \begin{equation} -->

<!-- f(y_i;\pi_i,\mu_i)=\begin{cases}\pi_i+(1-\pi_i)f(0;\mu_i)&y_i=0\\ -->

<!-- (1-\pi_i)f(y_i;\mu_i)&y_i>0\\ -->

<!-- \end{cases} -->

<!-- (\#eq:zi) -->

<!-- \end{equation} -->

<!-- The two link functions are -->

<!-- \begin{equation} -->

<!-- \begin{split} -->

<!-- g_0(\pi_i)=&\mathbf{w}'_i\boldsymbol\gamma\\ -->

<!-- g_1(\mu_i)=&\mathbf{x}'_i\boldsymbol\beta\\ -->

<!-- \end{split} -->

<!-- \end{equation} -->

<!-- Note that the two mechanisms could have different covariates and coefficients.  -->

<!-- But both of $\pi_i$ and $\mu_i$ appear in two equations and have to be evaluated jointly. Newton-Raphson algorithm or EM algorithm can deal with this question. -->

<!-- Hurdle models is another type of two-step models. -->

<!-- Hurdle models assume that all zero observations are generated by the first mechanism. Hence the first mechanism is not depend on $\mathbf{x}_i$ and $\boldsymbol\beta$.  -->

<!-- A challenge is that ordinary Poisson or negative binomial distribution does contain zero values. Here use a truncated distribution to address this issue. -->

<!-- \begin{equation} -->

<!-- f(y_i;\pi_i,\mu_i)=\begin{cases}\pi_i&y_i=0\\ -->

<!-- (1-\pi_i)\frac{f(y;\mu_i)}{1-f(0;\mu_i)}&y_i>0\\ -->

<!-- \end{cases} -->

<!-- (\#eq:hurdle) -->

<!-- \end{equation} -->

<!-- where $f(0;\mu_i)= \exp[-\mu_i]$ in Poisson model and  $f(0;\mu_i)= (\frac{r}{\mu_i+r})^r$ in negative binomial model. -->

### Discrete Choice Models

<!--   Mode Choice-->

When the response is categorical variable, such as binary data of driving versus not, polytomous data of mode choice, or count data of trip frequency, the proper model is the expected response variable equals a function on the linear combination of coefficients and covariates. This function is called link function. About half of the 200 studies of transit use choose this framework [@astonExploringBuiltEnvironment2021]. <!-- \begin{equation} --> <!-- E[\mathbf{Y}|\mathbf{X}]=\mu=g^{-1}(\mathbf{X}\boldsymbol{\beta}) --> <!-- (\#eq:glm) --> <!-- \end{equation} -->

<!-- Where $E[\mathbf{Y}|\mathbf{X}]$ is the expected value of travel choice variable $\mathbf{Y}$ conditional on $\mathbf{X}$; $\mathbf{X}\boldsymbol{\beta}$ is a linear combination (Ditto.); $g(\cdot)$ is a link function. -->

Mode choice is the classical topic in travel studies. One can choose whether or not to take a trip, driving or active modes, which is the binary data. For travel mode choice, a traveler has more than two alternatives: driving, transit, biking, and walking, which is called polychotomous data. These discrete response variables cannot be denoted by continuous variables and can not use regular linear models. Logistic regression models can address this problem. @mcfaddenConditionalLogitAnalysis1973 proposed the Discrete Choice Models, also called multinomial/conditional logit model. The detailed technique can be found in @agresti2015foundations and @rodriguez2007generalized. The further discussion of the limitation of logistic models can be found in @lipseyPracticalMetaanalysis2001, chap. 3. Multinomial logistic regression models are widely used in mode choice/split questions. Among the 45 selected studies of mode choice, most of them use the logistic regression models.

An alternative is the probit model. A recent application can be found in @chakourExaminingInfluenceStop2016. They use the ordered probit (ORP) model to examine the effects of the transit system and built environment on the probability of ridership.

<!-- witch assumes the error terms $\boldsymbol\varepsilon\sim MVN(\mathbf{0},\Sigma)$ where $\Sigma$ is a correlation matrix.  -->

<!-- Logistic models are not robust when the probability of $\pi$ is close to zero or one.  -->

<!-- For mode choice questions, the proportions of walking, biking, and transit are much smaller than that of driving. -->

<!-- In logistic models, the goal is to estimate the unknown vector of parameters $\boldsymbol\beta$ for the known covariates $\mathbf{X}_i$.  -->

<!-- But in the systematic component, $\eta_i$ is unobserved. Ordinary Linear Regression doesn't work in this case.  -->

<!-- Fortunately, the link function in logit models has a close form. -->

<!-- Iteratively Reweighted Least Squares method (IRLS) (Lawson 1961) can get the solution.  -->

<!-- When the probability and sample size of one mode is small (e.g. $\hat\pi_i=0.05$), IRLS would be assigned a small weight. "The standard error is artificially compressed, which leads us to overestimate the precision of the proportion estimate." [@lipseyPracticalMetaanalysis2001, chap. 3] -->

<!-- Sometimes, researcher can combine several modes such as walking and biking to active mode and relief this issue. -->

<!-- Otherwise, one has to looking for other algorithm, such as data augmentation by Markov chain Monte Carlo (MCMC) to get the more stable estimates. -->

<!-- Generalized Linear Models (GLM) allow the response following more general distributions than normal. -->

<!-- GLMs include three components.  -->

<!-- Systematic component $\eta=\mathbf{X}\boldsymbol\beta$ has a similar form with ordinary linear models but without error term. $\boldsymbol\beta$ are unknown coefficients. -->

<!-- Random component $E[Y]=\mu$ specifies the probability distributions of $Y$, which could have a pdf or pmf from an exponential family. -->

<!-- Link function $g(\cdot)$ connects the systematic component and random component together. -->

<!-- (equation @ref(eq:glm))  -->

<!-- - Binomial Response -->

<!-- When a traveler choose to make a trip or not, the decision follows a Bernoulli distribution. The probability is denoted by $Pr(\text{choice}=\text{Yes})=\pi$ and $Pr(\text{choice}=\text{No})=1-\pi$.  -->

<!-- For $n$ number of decisions under the same $\pi$, let $Y$ represents the count of choosing 'Yes' and follow a binomial distribution $Bin(n,\pi)$. -->

<!-- For many travelers with different $\pi$, one has $Y_i\sim Bin(n_i,\pi_i)$, that is a binary response data. The number of total observation $N=\sum_{i=1}^n n_i$. The pmf of binomial distribution is -->

<!-- \begin{equation} -->

<!-- Pr(Y_i = y_i) =  {{n_i}\choose{y_i}}  \pi_i^{y_i} (1-\pi_i)^{n_i-y_i} -->

<!-- \end{equation} -->

<!-- It is clear that the random component is $E[y_i]=\pi_i$ and systematic component $\eta_i=\mathbf{X}'_i\boldsymbol\beta$.  -->

<!-- $\pi$ is the probability between zero and one. but the log odds of success $\eta_i$ can take any real number. The canonical form of binomial distribution is -->

<!-- \begin{equation} -->

<!-- Pr(Y_i = y_i) = \exp\left[\log(\frac{\pi_i}{1-\pi_i})y_i+n_i\log(1-\pi_i)\right]{{n_i}\choose{y_i}} -->

<!-- \end{equation} -->

<!-- The canonical link function in logit models can transform the probability to the range of real number. In this one-to-one mapping, a probability $\pi_i>1/2$ will give a positive $\eta_i$ and a negative $\eta_i$ correspond to a $\pi_i$ less than one half. -->

<!-- \begin{equation} -->

<!-- \begin{split} -->

<!-- g(\pi_i)&=\log\frac{\pi_i}{1-\pi_i}=\eta_i\quad\text{Logit function}\\ -->

<!-- g^{-1}(\eta_i)&=\frac{\exp[\eta_i]}{1+\exp[\eta_i]}=\pi_i\quad\text{Logistic function}\\ -->

<!-- \end{split} -->

<!-- (\#eq:logit-link) -->

<!-- \end{equation} -->

<!-- ### Probit Models (Opt.) -->

<!-- - Multinomial Response -->

<!-- For categorical response such as travel mode choice, a traveler has more than two alternatives including driving, transit, biking and walking. The generalized logistic regression can address these polychotomous data. -->

<!-- The mode choice $Y_i$ follows the mutinomial distribution with $J$ alternatives.  -->

<!-- Denote the probability of $i$th traveler chooses the $j$th mode, then $\pi_{ij}=Pr(Y_i=j)$. -->

<!-- And the pmf of multinomial distribution is  -->

<!-- \begin{equation} -->

<!-- Pr(Y_{i1}=y_{i1}, ..., Y_{iJ}=y_{iJ})= {n_i \choose y_{i1},..., y_{iJ} } -->

<!-- \pi_{i1}^{y_{i1}} \cdots \pi_{iJ}^{y_{iJ}} -->

<!-- \end{equation} -->

<!-- When the data exclude the people without trip, the several modes exhaust all observations and mutually exclusive. That is $\sum_{j=1}^J\pi_{ij}=1$ for each $i$. Once $J-1$ parameters are evaluated, the rest one will be determined automatically. That means $\pi_{iJ}=1-\pi_{i1}-\cdots-\pi_{i,J-1}$.  -->

<!-- The random component is $\mu_i=n_i\pi_{ij}$ and the systematic component is $\eta_{ij}=\mathbf{X}_i'\boldsymbol\beta_j$ -->

<!-- \begin{equation} -->

<!-- \begin{split} -->

<!-- g^{-1}(\eta_{ij})&=\frac{\exp[\eta_{ij}]}{\sum_{k=1}^J\exp[\eta_{ik}]}=\pi_{ij}\\ -->

<!-- g(\pi_{ij})&=\log\frac{\pi_{ij}}{\pi_{iJ}}=\eta_{ij}\\ -->

<!-- \end{split} -->

<!-- (\#eq:mlogit-link) -->

<!-- \end{equation} -->

<!-- @mcfaddenConditionalLogitAnalysis1973 proposed the Discrete Choice Models which is also called multinomial/conditional logit model. -->

<!-- This model introduces $U_{ij}$ as the random utility of $j$th choice. Then based on Utility Maximum theory,  -->

<!-- \begin{equation} -->

<!-- \pi_{ij}=Pr(Y_i=j)=Pr(\max(U_{i1},...,U_{iJ})=U_{ij}) -->

<!-- \end{equation} -->

<!-- Here $U_{ij}=\eta_{ij}+\varepsilon_{ij}$ where the error term follows a standard Type I extreme value distributions. The reason is that the difference between two independent extreme value distributions has a logistic distribution. Hence, it can still be solved by logit models. -->

<!-- The expected utility depend on the characteristics of the alternatives rather than that of individuals. -->

<!-- Let $\mathbf{Z}_j$ represents the characteristics of $j$th alternative, one has $\eta_{ij}=\mathbf{Z}_i'\boldsymbol\gamma$. -->

<!-- Combining the two sources of utility together, a general form of utility is $\eta_{ij}=\mathbf{X}_i'\boldsymbol\beta_j+\mathbf{Z}_i'\boldsymbol\gamma$ -->

<!-- One iteration of this approach includes four steps. -->

<!-- The first step starts from the current estimated coefficients $\boldsymbol{\hat\beta}^{(0)}$. Then the current estimates of $\hat{\eta_i}^{(0)}=\mathbf{x}_i'\boldsymbol{\hat\beta}^{(0)}$ and $\hat \pi_i^{(0)}=\frac{\exp[\hat\eta_i^{(0)}]}{1+\exp[\hat\eta_i^{(0)}]}$. -->

<!-- But current $\hat{\eta_i}^{(0)}$ is different with the true $\eta_i$. -->

<!-- The second step will update the current estimates by adding a correction term. -->

<!-- Using the first two terms of Taylor series, -->

<!-- \begin{equation} -->

<!-- \hat{\eta_i}^{(1)}=\hat{\eta_i}^{(0)} +(y_i-\hat\mu_i^{(0)})\cdot \frac{d\hat\eta_i^{(0)}}{d\hat\mu_i^{(0)}} -->

<!-- \end{equation} -->

<!--  Since $E[Y_i]=\mu_i=n_i\pi_i$, it is also easy to get  -->

<!-- \begin{equation} -->

<!-- \frac{d \eta_i}{d \mu_i}=\frac{1}{n_i}\cdot\frac{d \eta_i}{d \pi_i}=\frac{1}{n_i}\left(\frac{1}{\pi_i}+\frac{1}{1-\pi_i}\right)=\frac1{n_i\pi_i(1-\pi_i)} -->

<!-- \end{equation} -->

<!-- Therefore, -->

<!-- \begin{equation} -->

<!-- \hat{\eta_i}^{(1)}=\hat{\eta_i}^{(0)} + \frac{y_i-n_i\hat\pi_i^{(0)}}{n_i\hat\pi_i^{(0)}(1-\hat\pi_i^{(0)})} -->

<!-- \end{equation} -->

<!-- The third step is to calculate the diagonal weight matrix $\mathbf{W}$ in the Fisher scoring algorithm. -->

<!-- It is known that the binomial distribution has $Var[Y_i]=n_i\pi_i(1-\pi_i)$.  -->

<!-- \begin{equation} -->

<!-- w_{ii}=\left[Var[Y_i](\frac{d \eta_i}{d \mu_i})^2\right]^{-1}= -->

<!-- n_i\hat\pi_i(1-\hat\pi_i) -->

<!-- \end{equation} -->

<!-- The final step improves estimate of $\boldsymbol{\hat\beta}^{(1)}$ using the weighted least-squares method -->

<!-- \begin{equation} -->

<!-- \hat{\boldsymbol{\beta}}^{(1)}=\mathbf{(X'WX)^{-1}}\mathbf{X'W}\boldsymbol{\hat{\eta}}^{(1)} -->

<!-- \end{equation} -->

<!-- The four steps will repeat until the procedure convergence. And the result gives -->

<!-- \begin{equation} -->

<!-- Var[\hat{\boldsymbol{\beta}}]=\mathbf{(X'WX)^{-1}} -->

<!-- \end{equation} -->
