# (PART\*) Part II Regression Analysis {.unnumbered}

# Linear Models

This Chapter will not give a complete introduction of linear regression. It only focus on some critical issues which are related with Travel-Urban Form models. This chapter only discusses the VMT-urban form models because the response are continue variables in regular linear models. The part of mode choice model is placed in Chapter of Generalized linear models.

## Assumptions

### Additive and linearity

For linear models, the most important assumptions are the additive and linear relationship between the response and predictors. Gravity Law discloses that travel distance has a multiplicative (inverse) relationship with the 'masses' of two places. If the population size can be a representative of built environment, the additive relationship will not hold. Previous studies also shows that the effect sizes of built environment with respect of travel are small and complex. There is not sufficient evidence to support or against the linear hypothesis.

### Independent Identically Distributed (IID)

Another essential assumption is that random error are Independent Identically Distributed (IID). Random error is also called residual, which refer to the difference between observed $\mathbf{y}$ and fitted $\mathbf{\hat y}$. $\mathbf{\hat y}$ are the linear combinations of predictors $\mathbf{X}$. residuals represent the part can not be explained by the model.

$$\begin{equation}
\mathbf{e}=\mathbf{y}-\mathbf{\hat y}
(\#eq:residual)
\end{equation}$$

The expected value, the variances, and the covariances among the random errors are the first- and second-moment of residuals. 'Identical' means that random errors should have zero mean and constant variance. The homogeneity of variance is also called homoscedasticity.

$$\begin{equation}
E(\varepsilon) = 0 \\
Var(\varepsilon) = \sigma^2\\
(\#eq:identical)
\end{equation}$$

'Independent' requires the random errors are uncorrelated. That is

$$\begin{equation}
Cov[\varepsilon_i,\varepsilon_j] = 0,\quad i\neq j
(\#eq:indenpendent)
\end{equation}$$

Once the conditions of IID are satisfied, the Gauss - Markov theorem \@ref(thm:g-m) proves that least-square method could give the minimum-variance unbiased estimators (MVUE) or called the best linear unbiased estimators (BLUE). These conditions are not strict and make regression method widely applicable.

::: {#g-m .theorem name="Gauss - Markov theorem"}
For the regression model \@ref(eq:lm) with the assumptions $E(\varepsilon) = 0$, $Var(\varepsilon) = \sigma^2$, and uncorrelated errors, the least-squares estimators are unbiased and have minimum variance when compared with all other unbiased estimators that are linear combinations of the $y_i$. (Montgomery et al., 2021)

Another version is that: Under Models II - VII, if $\lambda'\beta$ is estimable and $\hat\beta$ is any solution to the normal equations, then $\lambda'\hat\beta$ is a linear unbiased estimator of $\lambda'\beta$ and, under Model II, the variance of $\lambda'\hat\beta$ is uniformly less than that of any other linear unbiased estimator of $\lambda'\beta$ (IX, Theorem E13, p38)
:::

Unfortunately, many of the predictors are correlated. Moreover, the observations from various cities, regions, or counties are very unlikely identical. This issue is called heteroscedasticity. Related contents are in Section of Diagonusis and Validation.

### Normality

When conducting hypothesis test and confidence intervals, the required assumption is $\mathbf{y|x}\sim N (\mathbf{X}\boldsymbol{\beta},\sigma^2\mathbf{I})$. Maximum Likelihood Methods also requires this assumption.

Evidence has demonstrated that travel distance is not Normal distributed. The Zipf's law also prove that travel distance follows a power distribution. Using logarithm transformations, the skewed distribution can be converted to an approximate normal distribution.

There are some quantitative methods which can examine nomalirty of the transformed distributions.

## Estimations

### Least Square

Least-Squares method can be used to estimate the coefficients $\beta$ in equation \@ref(eq:lm) The dimension of $\mathbf{X}$ is $n\times p$, which means the data contain $n$ observations and $p-1$ predictors. The $p\times1$ vector of least-squares estimators is denoted as $\hat\beta$ and the solution to the normal equations is

$$\begin{equation}
\boldsymbol{\hat\beta}=(\mathbf{X'X})^{-1}\mathbf{X'}\mathbf{y}
(\#eq:lsq)
\end{equation}$$

Here requires $\mathbf{X'X}$ are invertible, that is, the covariates are linearly independent if $\mathbf{X}$ has rank $p$ (V., Definition, p.22).

Given the estimated coefficients, the model can give the fitted values of response as:

$$\begin{equation}
\mathbf{\hat y}=\mathbf{X}\boldsymbol{\hat\beta}=\mathbf{X}(\mathbf{X'X})^{-1}\mathbf{X'y}= \mathbf{Hy}
(\#eq:fitted-y)
\end{equation}$$

where $\mathbf{H}=\mathbf{X}(\mathbf{X'X})^{-1}\mathbf{X'}$ is hat matrix and $\mathbf{e}=\mathbf{y}-\mathbf{\hat y}=\mathbf{y}-\mathbf{X}\boldsymbol{\hat\beta}=(\mathbf{I}-\mathbf{H})\mathbf{y}$

When the observations are not independent or have unequal variances, the covariance matrix of error is not identity matrix. The assumption of regression model $V[\boldsymbol{\varepsilon}]=\sigma^2\mathbf{I}$ doesn't hold. Denote $\mathbf{V}$ is a known $n\times n$ positive definite matrix and $V[\boldsymbol{\varepsilon}]=\sigma^2\mathbf{V}$.
Then, there exists an $n\times n$ symmetric matrix $\mathbf{K}$ with rank $n$ and $\mathbf{V}=\mathbf{KK'}$. Let

$$\begin{equation}
\mathbf{z}=\mathbf{K'y},\ \mathbf{B}=\mathbf{K^{-1}X}, \text{and}\ \boldsymbol{\eta}=\mathbf{K'}\boldsymbol{\varepsilon}
\end{equation}$$

The linear model becomes $\mathbf{z}=\mathbf{B}\boldsymbol{\beta}+\boldsymbol{\eta}$ and $V[\boldsymbol{\eta}]=\sigma^2\mathbf{I}$.
If the model is full rank, that is $rank(\mathbf{X})=p$ then $\mathbf{X'V^{-1}X}$ is invertible
and the generalized least squares solution is

$$\begin{equation}
\boldsymbol{\hat\beta}_{GLS}=(\mathbf{X'V^{-1}X})^{-1}\mathbf{X'V^{-1}}\mathbf{y}
(\#eq:glsq)
\end{equation}$$



### Standardized coefficients

The value of $\hat \beta_j$ means that, given all other coefficients fixed, for each change of one unit in $x_j$, the average change in the mean of $Y$. However, the units of predictors $\mathbf{X}$ are very different. Hence, the values of coefficients are not comparable.

Unit normal scaling or Unit length scaling can convert $\hat \beta_j$ to dimensionless regression coefficient, which is called standardized regression coefficients. Let

$$\begin{equation}
\begin{split}
z_{ij}=&\frac{x_{ij}-\bar x_j}{\sqrt{\sum_{i=1}^{n}(x_{ij}-\bar x_j)^2}},\quad \\
y^{0}_{i}=&\frac{y_{i}-\bar y}{\sqrt{\sum_{i=1}^{n}(y_{i}-\bar y)^2}}
\end{split}
(\#eq:standize)
\end{equation}$$

$$\begin{equation}
\begin{split}
\mathbf{\hat b}=&(\mathbf{Z'Z})^{-1}\mathbf{Z'}\mathbf{y^{0}},\ \text{or}\\
\hat b_j= &\hat\beta_j\sqrt{\frac{\sum_{i=1}^{n}(x_{ij}-\bar x_j)^2}{\sum_{i=1}^{n}(y_{i}-\bar y)^2}},\ j=1,2,...(p-1),\text{ and}\\
\hat\beta_0=&\bar y - \sum_{j=1}^{p-1}\hat\beta_j\bar x_j
\end{split}
(\#eq:stand-coef)
\end{equation}$$

Note that $\mathbf{Z'Z}$ correlations matrix.

$$\begin{equation}
\begin{split}
\mathbf{Z'Z}=&\begin{bmatrix} 
1 & r_{12} & r_{13} & \dots & r_{1k} \\  
r_{21} & 1 & r_{23} & \dots & r_{2k} \\  
r_{31} & _{32} & 1 & \dots & r_{3k} \\  
\vdots & \vdots & \vdots & \ddots & \vdots \\  
r_{k1} & r_{k2} & _{k3} & \dots & 1  \end{bmatrix},\quad 
\mathbf{Z'}\mathbf{y^{0}}=&\begin{bmatrix} 
r_{1y} \\ r_{2y} \\ r_{3y} \\ \vdots \\ r_{ky} \end{bmatrix}
\end{split}
(\#eq:corr-matrix)
\end{equation}$$

where

$$\begin{equation}
\begin{split}
r_{ij}=&\frac{\sum_{u=1}^{n}(x_{ui}-\bar x_i)(x_{uj}-\bar x_j)}{\sqrt{\sum_{u=1}^{n}(x_{ui}-\bar x_i)^2\sum_{u=1}^{n}(x_{uj}-\bar x_j)^2}}\\
r_{jy}=&\frac{\sum_{u=1}^{n}(x_{uj}-\bar x_j)(y_{u}-\bar y)}{\sqrt{\sum_{u=1}^{n}(x_{uj}-\bar x_j)^2\sum_{u=1}^{n}(y_{u}-\bar y)^2}}
\end{split}
(\#eq:corr-1)
\end{equation}$$

where $r_{ij}$ is the simple correlation between $x_i$ and $x_j$. $r_{jy}$ is the simple correlation between $x_j$ and $y$

It seems that standardized regression coefficients are comparable. However, the value of $\hat b_j$ depends on other predictors. Therefore, comparison between different models is still problematic.

### Elasticity

Definition: Commonly used to determine the relative importance of a variable in terms of its influence on a dependent variable. It is generally interpreted as the percent change in the dependent variable induced by a 1% change in the independent variable.[^linear-1]

$$e_i=\beta_i\frac{X_i}{Y_i}\approx\frac{\partial Y_i}{\partial X_i}\frac{X_i}{Y_i}$$

[^linear-1]: McCarthy, P.S., Transportation Economics Theory and Practice: A Case Study Approach. Blackwell, Boston, 2001.

+----------------+-------------------------------+---------------------------------+
| Model          | Marginal Effects              | Elasticity                      |
+:==============:+:=============================:+:===============================:+
| Linear         | $\beta$                       | $\beta\frac{X_i}{Y_i}$          |
+----------------+-------------------------------+---------------------------------+
| Log-linear     | $\beta Y_i$                   | $\beta X_i$                     |
+----------------+-------------------------------+---------------------------------+
| Linear-log     | $\beta\frac{1}{X_i}$          | $\beta\frac{1}{Y_i}$            |
+----------------+-------------------------------+---------------------------------+
| Log-log        | $\beta\frac{Y_i}{X_i}$        | $\beta$                         |
+----------------+-------------------------------+---------------------------------+
| Logit          | $\beta p_i(1-p_i)$            | $\beta X_i(1-p_i)$              |
+----------------+-------------------------------+---------------------------------+
| Poisson        | $\beta \lambda_i$             | $\beta X_i$                     |
+----------------+-------------------------------+---------------------------------+
| NB             | $\beta \lambda_i$             | $\beta X_i$                     |
+----------------+-------------------------------+---------------------------------+

: Elasticity Estimates for Various Functional Forms

It is strange that @ewingTravelBuiltEnvironment2010 use the formula of $\beta \bar X\left(1-\frac{\bar Y}{n}\right)$ for Logit model.
In Poisson model and Negative Binomial model, $\lambda_i=\exp[\mathbf{x}_i'\boldsymbol{\beta}]$ (Greene, 2018, eq.18-17,21).
For truncated Poisson model: $\delta_i=\frac{(1-P_{i,0}-\lambda_i P_{i,0})}{(1-P_{i,0})^2}\cdot\lambda_i\beta$ (Greene, 2018, eq.18-23).
Hurdle model will give separate marginal(partial) effects (Greene, 2018, example 18.20)

### Combined effects?

Some studies sums up the standardized coefficients or elasticities and called the summation as combined effects. Although these values are dimensionless, this method is problematic because different model specifications and data ranges are not comparable.

## Inference

### Analysis of Variance

Analysis of Variance (ANOVA) is the fundamental approach in regression analysis. Actually, this method analysis the variation in means rather than variances themselves (Casella & Berger, 2002, Ch.11).

Once the linear relationship holds, the response $\mathbf{y}$ can be decomposed to

$$\begin{equation}
\begin{split}
\mathbf{y'y}=&\mathbf{y'Hy}+\mathbf{y'}(\mathbf{I}-\mathbf{H})\mathbf{y}\\
\mathbf{y'y}=&\boldsymbol{\hat\beta}\mathbf{X'}\mathbf{y}+\mathbf{y'y}-\boldsymbol{\hat\beta}\mathbf{X'}\mathbf{y}\\
\mathbf{y'y}-n\bar y^2=&\boldsymbol{\hat\beta}\mathbf{X'}\mathbf{y}-n\bar y^2+\mathbf{y'y}-\boldsymbol{\hat\beta}\mathbf{X'}\mathbf{y}\\
\sum(y-\bar y)^2=&\sum(\hat y-\bar y)^2+\sum(y-\hat y)^2\\
SST =& SSR + SSE
\end{split}
(\#eq:ss)
\end{equation}$$

where $SST$ is Sum of Squares Total, $SSR$ is Sum of Squares Regression, and $SSE$ is Sum of Square Error. $SSE=\mathbf{e'e}$ represents the unknown part of model.

For Generalized Least Squares method, $SST=\mathbf{y'V^{-1}y}$, $SSR= \boldsymbol{\hat\beta'}\mathbf{B'z}=\mathbf{y'V^{-1}(X'V^{-1}X})^{-1}\mathbf{X'V^{-1}}\mathbf{y}$, and $SSE=SST-SSR$

### Hypothesis Test

-   Significance of Regression

Significance of regression means if the linear relationship between response and predictors is adequate. The hypotheses for testing model adequacy are

$$\begin{equation}
\begin{split}
H_0:&\quad \beta_0 = \beta_1 = \cdots =\beta_{p-1}=0\\
H_1:&\quad \text{at least one } \beta_j \neq 0,\ j=0,1,...,(p-1)\\
\end{split}
(\#eq:hyp-1)
\end{equation}$$

By Theorem D14 (XX, p.90), if an $n\times1$random vector $\mathbf{y}\sim N(\boldsymbol{\mu},\mathbf{I})$, then

$$\begin{equation}
\mathbf{y'y} \sim \chi^2(n,\frac12\boldsymbol{\mu'\mu})
(\#eq:chisq)
\end{equation}$$

Recall the assumption of $\mathbf{y|x}\sim N (\mathbf{X}\boldsymbol{\beta},\sigma^2\mathbf{I})$.\
By the additive property of $\chi^2$ distribution,

$$\begin{equation}
\frac{MSE}{\sigma^2}=\frac{\mathbf{y'(I-H)y}}{(n-p)\sigma^2} \sim \chi^2_{(n-p)}\\
\frac{MSR}{\sigma^2}=\frac{\mathbf{y'Hy}}{(p-1)\sigma^2} \sim \chi^2_{(p-1)}\\
\end{equation}$$

Though $\sigma^2$ is usually unknown, by the relationship between $\chi^2$ and $F$ distributions,

$$\begin{equation}
F_0=\frac{MSE}{MSR} \sim F_{(p-1),(n-p),\lambda}\\
\end{equation}$$ where $\lambda$ is the non-centrality parameter. It allows to test the hypotheses given a significance level $\alpha$. If test statistic $F_0>F_{\alpha,(p-1),(n-p)}$, then one can reject $H_0$.

If a VMT-urban form model added many predictors but adjusted $R^2$ is still low, the association between travel distance and built environment might be spurious.

-   Significance of Coefficients

For testing a specific coefficient, the hypothesis is

$$\begin{equation}
\begin{split}
H_0:&\quad \beta_j =0\\
H_1:&\quad \beta_j \neq 0\\
\end{split}
(\#eq:hyp-2)
\end{equation}$$

$\boldsymbol{\hat\beta}$ is a linear combination of $\mathbf{y}$. Based on the assumption of $\mathbf{y|x}\sim N (\mathbf{X}\boldsymbol{\beta},\sigma^2\mathbf{I})$, it can be proved that $\boldsymbol{\hat\beta}\sim N (\boldsymbol{\beta},\sigma^2(\mathbf{X'X})^{-1})$ and

$$\begin{equation}
t_0=\frac{\hat\beta_j}{se(\hat\beta_j)}=\frac{\hat\beta_j}{\sqrt{\hat\sigma^2C_{jj}}} \sim t_{(n-p)}\\
\end{equation}$$ where $C_{jj}$ is the element at the $j$ row and $j$ column of $(\mathbf{X'X})^{-1}$. If $|t_0|< t_{\alpha/2,(n-p)}$, then the test failed to reject the $H_0$, this predictor can be removed from the model. This test is called partial or marginal test because the test statistic for $\beta_j$ depends on all the predictors in the model.

### Confidence Intervals

Above results can also construct the confidence interval for each coefficient. A $100(1-\alpha)$ confidence interval for $\beta_j$ is

$$\begin{equation}
\hat\beta_j-t_{\alpha/2,(n-p)}\sqrt{\hat\sigma^2C_{jj}}\le \beta_j \le \hat\beta_j+t_{\alpha/2,(n-p)}\sqrt{\hat\sigma^2C_{jj}}
\end{equation}$$

## Adequacy

Only estimation and inference can not demonstrate model's performance.
If the primary assumptions is violated, the estimations could be biased and the model could be useless. These problems can also happen when the model is not correctly specified. It is necessary to make diagnosis and validation for fitted models.



<!-- "There are different types of model specification errors in regression analysis: " -->
<!-- (1) necessary independent variable(s) is (are) missing from the model (i.e., omission error) (Greene 2012, 96--8, Section 4.3.2); -->
<!-- (2) irrelevant independent variable(s) is (are) added to the model (i.e., commission error) (Greene 2012, 98, Section 4.3.3); -->
<!-- (3) independent variables have an incorrect functional form (Ramsey 1969; Thursby and Schmidt 1977; Hausman 1978; Davidson and MacKinnon 1981); -->
<!-- (4) variables in the model are not accurately measured (i.e., measurement error) (Wooldridge 2015, 287--92, Section 9.4); and -->
<!-- (5) the model is not stand-alone but instead belongs to a system of simultaneous equations (Ramsey 1969; Hausman 1978). -->

### Residuals Analysis

The major assumptions, both IID and normality are related to residual. 
Residual diagnosis is an essential step for modeling validation.

There are several scaled residuals can help the diagnosis. 
Since $MSE$ is the expected variance of error $\hat\sigma^2$ and $E[\varepsilon]=0$, standardized residuals should follow a standard normal distribution.

$$d_i=\frac{e_i}{\sqrt{MSE}}=e_i\sqrt{\frac{n-p}{\sum_{i=1}^n e_i^2}},\quad i=1,2,...,n$$

Recall random error $\mathbf{e}=\mathbf{y}-\mathbf{\hat y}=(\mathbf{I}-\mathbf{H})\mathbf{y}$ and hat matrix $\mathbf{H}=\mathbf{X}(\mathbf{X'X})^{-1}\mathbf{X'}$.
Let $h_{ii}$ denote the $i^{th}$ diagonal element of hat matrix.
Studentized Residuals can be expressed by
$$r_i=\frac{e_i}{\sqrt{MSE(1-h_{ii})}},\quad i=1,2,...,n$$
It is proved that $0\le h_{ii}\le1$.
An observation with $h_{ii}$ closed to one will return a large value of $r_i$. The $x_i$ who has strong influence on fitted value is called leverage point.

Ideally, the scaled residual have zero mean and unit variance. Hence, an observation with $|d_i|>3$ or $|r_i|>3$ is a potential outlier.

Predicted Residual Error Sum of Squares (PRESS) can also be used to detect outliers.
This method predicts the $i^{th}$ fitted response by excluding the $i^{th}$ observation and examine the influence of this point.
The corresponding error $e_{(i)}=e_{i}/(1-h_{ii})$ and $V[e_{(i)}]=\sigma^2/(1-h_{ii})$.
Thus, if $MSE$ is a good estimate of $\sigma^2$, PRESS residuals is equivalent to Studentized Residuals.

$$\frac{e_{(i)}}{\sqrt{V[e_{(i)}]}}=\frac{e_i/(1-h_{ii})}{\sqrt{\sigma^2/(1-h_{ii})}}=\frac{e_i}{\sqrt{\sigma^2(1-h_{ii})}}$$

- Residual Plot

Residual plot shows the pattern of the residuals against fitted $\mathbf{\hat y}$.
If the assumptions are valid, the shape of points should like a envelope and be evenly distributed around the horizontal line of $e=0$.

A funnel shape in residual plot shows that the variance of error is a function of $\hat y$. A suitable transformation to response or predictor could stabilize the variance.
A curved shape means the assumption of linearity is not valid. It implies that adding quadratic terms or higher-order terms might be suitable.


- Normal Probability Plot

A histogram of residuals can check the normality assumption. 
The highly right-skewed probability distribution of VMT log-transform of VMT is reasonable.

A better way is a normal quantile -- quantile (QQ) plot of the residuals.
An ideal cumulative normal distribution should plot as a straight line.
Only looking at the $R^2$ and p-values cannot disclose this feature. 


### Heteroscedasticity




Bartlett's (1937) likelihood ratio test (Ravishanker & Dey, 2020, 8.1.3, p.288)

### Multicollinearity

Multicollinearity or near-linear dependence refers to the models with highly correlated predictors. When data is generated from experimental design, the treatments $X$ could be fixed variables and be orthogonal. But travel-urban form model is observational studies and nothing can be controlled as in lab. It is known that there are complex correlations among the built-environment predictors themselves.

Although, the basic IID assumptions do not require that all predictors $\mathbf{X}$ are independent, when the predictors are near-linear dependent, the model is ill-conditioned and the least-square estimators are unstable.

multicollinearity can make the variances inflated and impact model precision seriously. If some of predictors are exact linear dependent, the matrix $\mathbf{X'X})^{-1}$ will be non-invertible. If the predictors are near-linear dependent, the correlation matrix using unit length scaling $\mathbf{Z'Z}$will has a inverse matrix with inflated variances. That means that the diagonal elements of $(\mathbf{Z'Z})^{-1}$ are not all equal to one. The diagnoal elements are called **Variance Inflation Factors**, which can be used to examine multicollinearity. The VIF for a particular predictor is examined as below \@ref(eq:vif)

$$\begin{equation}
VIF_j=\frac{1}{1-R_j^2}
(\#eq:vif)
\end{equation}$$

where $R_j^2$ is the coefficient of determination by regressing $x_j$ on all the remaining predictors.

A common approach is to drop off the predictor with greatest VIF and refit the model until all VIFs are less than 10. However, dropping off one or more predictors will lose many information which might be valuable for explaining response. Due to the complexity among predictors, dropping off the predictor with the greatest VIF is not always the best choice. Sometimes, removing a predictor with moderate VIF can make all VIFs less than 10 in the refitted model. Moreover, there is not an unique criteria for VIF value. When the relationship between predictor and response is weak, or the $R^2$ is low, the VIFs less than 10 may also affect the ability of estimation dramatically.

Orthogonalization before fitting the model might be helpful. Other approaches such as ridge regression or principal components regression could deal with multicollinearity better.

-   Subset Selection

Empirical or methodological considerations

-   Shrinkage Methods

#### Ridge Regression and Lasso

-   Dimention Reduction Methods

#### Principal Components Regression (PCR)

Partial Least Square (PLS)

By dint of some synthetic variables, the disaggregated model's $R^2$ can be over 0.5. But the risk is these techniques may describe the data themselves, and the results cannot be generalized. It is worthy of a deeper investigation.

```{r,eval=F}
# Compute k-means with k = 3
set.seed(123)
res.km <- kmeans(scale(iris[, -5]), 3, nstart = 25)
# K-means clusters showing the group of each individuals
fviz_cluster(res.km, data = iris[, -5],
             palette = "set3", #c("#2E9FDF", "#00AFBB", "#E7B800")
             geom = "point", pointsize=1, # "text"
             ellipse = T,ellipse.type ="t",   #'confidence'"euclid""convex"
             ellipse.level = 0.75, ellipse.alpha = 0.1,
             ggtheme = theme_bw(), main="Models with trip response variable",xlab="data", ylab="models"
             )
```

```{r,eval=F}
res.km <- kmeans(scale(USArrests), 4, nstart = 25)
fviz_cluster(res.km, data = USArrests,
             palette = "set3", #c("#2E9FDF", "#00AFBB", "#E7B800")
             geom = "point", pointsize=1, # "text"
             ellipse = T,ellipse.type ="t",   #'confidence'"euclid""convex"
             ellipse.level = 0.75, ellipse.alpha = 0.1,legend = "none",
             ggtheme = theme_bw(), main="Models with trip response variable",xlab="data", ylab="models"
             )
```

```{r modelcluster, eval=F, out.width='50%',fig.cap="A diagram of model cluster"}
# Dimension reduction using PCA
res.pca <- prcomp(USArrests,  scale = TRUE) #iris[, -5]
# Coordinates of individuals
ind.coord <- as.data.frame(get_pca_ind(res.pca)$coord)
# Add clusters obtained using the K-means algorithm
res.km <- kmeans(scale(USArrests), 4, nstart = 25)
ind.coord$cluster <- factor(res.km$cluster)
# Add Species groups from the original data sett
# ind.coord$Species <- iris$Species
eigenvalue <- round(get_eigenvalue(res.pca), 1)
variance.percent <- eigenvalue$variance.percent
p <- ggscatter(ind.coord,x = "Dim.1", y = "Dim.2",
  color = "cluster", palette = "npg", title="Find some proper Model clusters",
  ellipse = TRUE, ellipse.type = "t", #"convex"
  ellipse.border.remove=T, ellipse.level = 0.85,
  shape = "cluster", #"Species"
  size = 1.5,legend = "none", ggtheme = theme_bw(),
  xlab = "Various scale or scope",ylab = "Various probability distribution",
) + # theme(axis.text = element_blank()) +
  stat_mean(aes(color = cluster), size = 8,shape = c(1,2,0,3))

ggpar(p,tickslab=F,ticks=F)
```

## Variables Selections

### Goodness of fit

This structure tell us how good the model can explain the data. Coefficient of Determination $R^2$ is a proportion to assess the quality of fitted model.

$$\begin{equation}
R^2 =\frac{SSR}{SST}=1-\frac{SSE}{SST}
(\#eq:rsq)
\end{equation}$$

when $R^2$ is close to $1$, the most of variation in response can be explained by the fitted model. Although $R^2$ is not the only criteria of a good model, it is often available in most published papers. Recall the discussion in Part I, the aggregated data will eliminate the difference among individuals, households, or neighborhoods. In the new variance structure, $SSE$ will be much less than disaggregated model. The $R^2$ in many disaggregate studies are around 0.3, while the $R^2$ in some aggregate studies can reach 0.8. A seriously underfitting model's outputs could be biased and unstable.

A fact is that adding predictors into the model will never decrease $R^2$. Thus the models with different number of predictors is not comparable. Adjusted $R^2$ can address this issue by introducing degree of freedom. The degree of freedom denotes the amount of information required to know.

$$\begin{equation}
\begin{split}
df_T =& df_R + df_E\\
n-1=&(p-1)+(n-p)
\end{split}
(\#eq:df)
\end{equation}$$

Then, the mean square (MS) of each sum of squares (SS) can be calculated by $MS=SS/df$. The mean square error $MSE$ is also called as the expected value of error variance $\hat\sigma^2=MSE=SSE/(n-p)$. $n-p$ is the degree of freedom. Then adjusted $R^2$ is

$$\begin{equation}
R_{adj}^2 = 1-\frac{MSE}{MST} = 1-\frac{SSE/(n-p)}{SST/(n-1)}
(\#eq:rsq-adj)
\end{equation}$$

Another similar method is $R^2$ for prediction based on PRESS.
Recall the PRESS statistic is the prediction error sum of square by fitting a model with $n-1$ observations.

$$\begin{equation}
PRESS = \sum_{i=1}^n(y_i-\hat y_{(i)})^2= \sum_{i=1}^n\left(\frac{e_i}{1-h_{ii}}\right)^2
(\#eq:press)
\end{equation}$$


A model with smaller PRESS has a better ability of prediction. The $R^2$ for prediction is

$$\begin{equation}
R_{adj}^2 = 1-\frac{PRESS}{MST}
(\#eq:rsq-pred)
\end{equation}$$



### Mallows $C_p$

### PRESS

-   Underfitting models

-   Overfitting models

-   Interaction Effects

-   Polynomial Regression Models



-   Theoretical considerations

## Fixed and Mixed Effects

There are two different settings in experimental design. One assigns the treatments having fixed effects on the response. The other assign the treatments having random effects on the response. In fixed-effect model, the interest is the effects of levels themselves. The levels of the factor are chosen to test the differences among these specific levels. The chosen levels exhaust the pop The fixed effects across cases are constant.

While in random-effect model, the interest is the underlying population. The random effects are vary.

In random-effect models, the levels are selected at random from a large number of levels.

In VMT-urban form models, the interest is the effects of built-environment variables, not the specific region's impact.

## Bayesian approaches (Opt.)

\pagebreak

## SEM (Opt.)

Another attempt tries the method of structural equation modeling (SEM). The two studies capture higher elasticities of per capita VMT with respect to density (-0.38 and -0.238) [@cerveroEffectsBuiltEnvironments2010; @ewingStructuralEquationModels2014].


In general, modeling is a case-by-case work. Researchers may have their preferred model by weighing the sensitivity and robustness even given the same hypothesis and data.
The published papers usually don't show the results of diagnosis and validation.
Under this circumstance, compare or summarize these outcomes are unreliable.
